---
title: "Medulloblastoma Liquid Biopsy Report"
subtitle: "CSF cfDNA Analysis"
output: 
  pdf_document:
    toc: false
params:
  sample_id: "MB001"
  classification_file: "classification.csv"
  risk_file: "risk.csv"
  ichorcna_dir: "ichorcna/"
  seg_file: "segs.txt"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
library(data.table)
library(ggplot2)
library(gridExtra)
library(knitr)
```

# EXECUTIVE SUMMARY

```{r load_data}
class_data <- fread(params$classification_file)
risk_data <- fread(params$risk_file)
seg_data <- fread(params$seg_file)
```

**Patient Sample:** `r params$sample_id`  
**Report Date:** `r format(Sys.Date(), "%B %d, %Y")`  
**Analysis:** CSF Cell-Free DNA (Low-Coverage WGS)

---

## Primary Findings

**Medulloblastoma Subgroup:** **`r class_data$MB_Subgroup`**

**Classification Confidence:** `r class_data$Confidence`

**Risk Stratification:** `r risk_data$Risk_Group`

**Expected 5-Year PFS:** `r risk_data$PFS_5year`

**Tumor Fraction (CSF):** `r round(class_data$Tumor_Fraction * 100, 2)`%

---

# MOLECULAR CLASSIFICATION

## Subgroup Assignment

Based on copy number alteration patterns detected in cerebrospinal fluid cell-free DNA, this sample is classified as **`r class_data$MB_Subgroup` medulloblastoma**.

```{r subgroup_scores}
scores <- data.frame(
  Subgroup = c("WNT", "SHH", "Group 3", "Group 4"),
  Score = c(class_data$WNT_Score, class_data$SHH_Score, 
            class_data$Group3_Score, class_data$Group4_Score)
)

ggplot(scores, aes(x=reorder(Subgroup, -Score), y=Score, fill=Subgroup)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=Score), vjust=-0.5) +
  scale_fill_manual(values=c("WNT"="#2E7D32", "SHH"="#1976D2", 
                             "Group 3"="#C62828", "Group 4"="#F57C00")) +
  labs(title="Subgroup Classification Scores",
       x="MB Subgroup", y="Score") +
  theme_minimal() +
  theme(legend.position="none")
```

### Detected Molecular Features

```{r features_table}
features <- data.frame(
  Feature = character(),
  Status = character(),
  Clinical_Significance = character()
)

if (class_data$Monosomy_6) {
  features <- rbind(features, data.frame(
    Feature = "Monosomy 6",
    Status = "PRESENT",
    Clinical_Significance = "Diagnostic for WNT subgroup"
  ))
}

if (class_data$Iso17q) {
  features <- rbind(features, data.frame(
    Feature = "Isochromosome 17q",
    Status = "PRESENT",
    Clinical_Significance = "Common in Group 3/4"
  ))
}

if (class_data$MYC_amp) {
  features <- rbind(features, data.frame(
    Feature = "MYC amplification",
    Status = "PRESENT",
    Clinical_Significance = "HIGH RISK - Group 3 marker"
  ))
}

if (class_data$MYCN_amp) {
  features <- rbind(features, data.frame(
    Feature = "MYCN amplification",
    Status = "PRESENT",
    Clinical_Significance = "High risk - SHH/Group 4"
  ))
}

if (class_data$Chr8_loss) {
  features <- rbind(features, data.frame(
    Feature = "Chromosome 8 loss",
    Status = "PRESENT",
    Clinical_Significance = "Suggests Group 4"
  ))
}

if (nrow(features) > 0) {
  kable(features, caption="Key Molecular Alterations")
} else {
  cat("No high-confidence molecular features detected.\n")
  cat("This may indicate low tumor fraction or inconclusive pattern.\n")
}
```

---

# RISK STRATIFICATION

**Risk Group:** `r risk_data$Risk_Group`

**Clinical Interpretation:**  
`r risk_data$Rationale`

**Expected Outcome:**  
5-Year Progression-Free Survival: `r risk_data$PFS_5year`

### Risk Modifiers

```{r risk_modifiers}
risk_mods <- data.frame(
  Modifier = c("MYC amplification", "MYCN amplification", 
               "Chromosome 13 loss", "Metastasis at diagnosis",
               "Residual disease post-surgery"),
  Present = c(
    ifelse(risk_data$MYC_amplification, "YES", "No"),
    ifelse(risk_data$MYCN_amplification, "YES", "No"),
    ifelse(risk_data$Chr13_loss, "YES (protective)", "No"),
    ifelse(risk_data$Metastasis, "YES", "Unknown"),
    ifelse(risk_data$Residual_disease, "YES", "Unknown")
  ),
  Impact = c("Adverse", "Adverse", "Protective", "Adverse", "Adverse")
)

kable(risk_mods, caption="Risk Modifiers")
```

---

# COPY NUMBER PROFILE

```{r cnv_plot, fig.width=10, fig.height=4}
# Prepare genome-wide CNV plot
seg_data$chr_num <- as.numeric(gsub("X", "23", seg_data$chr))
seg_data <- seg_data[order(seg_data$chr_num, seg_data$start)]

# Chromosome positions
chr_info <- seg_data[, .(max_pos=max(end)), by=chr_num]
chr_info <- chr_info[order(chr_info$chr_num)]
chr_info$cum_pos <- cumsum(as.numeric(chr_info$max_pos))
chr_info$start_pos <- c(0, chr_info$cum_pos[-nrow(chr_info)])

seg_data <- merge(seg_data, chr_info[, .(chr_num, start_pos)], by="chr_num")
seg_data$plot_start <- seg_data$start_pos + seg_data$start
seg_data$plot_end <- seg_data$start_pos + seg_data$end
seg_data$plot_mid <- (seg_data$plot_start + seg_data$plot_end) / 2

ggplot(seg_data, aes(x=plot_mid, y=copy.number)) +
  geom_segment(aes(x=plot_start, xend=plot_end,
                   y=copy.number, yend=copy.number,
                   color=copy.number), size=2) +
  geom_hline(yintercept=2, linetype="dashed", color="gray40") +
  geom_hline(yintercept=c(1, 3), linetype="dotted", color="gray60") +
  scale_color_gradient2(low="blue", mid="gray80", high="red",
                       midpoint=2, name="Copy\nNumber") +
  scale_x_continuous(breaks=chr_info$start_pos + chr_info$max_pos/2,
                     labels=chr_info$chr_num) +
  scale_y_continuous(limits=c(0, max(5, max(seg_data$copy.number, na.rm=TRUE)))) +
  labs(title="Genome-Wide Copy Number Profile",
       x="Chromosome", y="Copy Number") +
  theme_minimal() +
  theme(panel.grid.minor=element_blank())
```

---

# QUALITY METRICS

```{r qc_table}
qc <- data.frame(
  Metric = c("Tumor Fraction", "Ploidy", "Classification Confidence", 
             "Number of CNV Segments"),
  Value = c(
    paste0(round(class_data$Tumor_Fraction * 100, 2), "%"),
    as.character(class_data$Ploidy),
    class_data$Confidence,
    nrow(seg_data)
  ),
  Status = c(
    ifelse(class_data$Tumor_Fraction >= 0.08, "PASS", "LOW"),
    "N/A",
    class_data$Confidence,
    "N/A"
  )
)

kable(qc, caption="Quality Control Metrics")
```

```{r qc_warning}
if (class_data$Tumor_Fraction < 0.08) {
  cat("\n**QUALITY ALERT:** Tumor fraction below recommended threshold (8%).\n")
  cat("Classification confidence may be reduced. Consider:\n")
  cat("- Repeat CSF sampling if clinically indicated\n")
  cat("- Correlation with tissue-based molecular testing\n")
  cat("- Serial monitoring to track tumor burden\n")
}
```

---

# CLINICAL INTERPRETATION

## Subgroup-Specific Considerations

```{r clinical_interp}
if (class_data$MB_Subgroup == "WNT") {
  cat("**WNT-activated medulloblastoma** has excellent prognosis (>90% 5-year survival).\n\n")
  cat("**Treatment Implications:**\n")
  cat("- Consider reduced-intensity therapy to minimize late effects\n")
  cat("- Gross total resection may not be necessary (survival excellent regardless)\n")
  cat("- Conservative surgical approach to reduce posterior fossa syndrome risk\n")
  cat("- Reduced craniospinal radiation dose may be appropriate\n")
  
} else if (class_data$MB_Subgroup == "SHH") {
  cat("**SHH-activated medulloblastoma** has variable prognosis depending on TP53 status.\n\n")
  cat("**Treatment Implications:**\n")
  cat("- Consider SMO inhibitors (vismodegib/sonidegib) for recurrent disease\n")
  cat("- TP53 mutation testing recommended for definitive risk stratification\n")
  cat("- MYCN amplification indicates poor prognosis\n")
  
} else if (class_data$MB_Subgroup == "Group3") {
  cat("**Group 3 medulloblastoma** has poor prognosis (<50% 5-year survival).\n\n")
  cat("**Treatment Implications:**\n")
  cat("- Aggressive multimodal therapy indicated\n")
  cat("- MYC amplification is very high risk - consider treatment intensification\n")
  cat("- Close surveillance for metastatic disease\n")
  cat("- Consider clinical trial enrollment\n")
  
} else if (class_data$MB_Subgroup == "Group4") {
  cat("**Group 4 medulloblastoma** has intermediate prognosis (~75% 5-year survival).\n\n")
  cat("**Treatment Implications:**\n")
  cat("- Standard risk-adapted therapy\n")
  cat("- Maximal safe resection improves PFS\n")
  cat("- Chromosome 13 loss is protective (if present)\n")
  cat("- MYCN amplification indicates higher risk\n")
}
```

---

# METHODS

**Sample Type:** Cerebrospinal fluid (CSF) cell-free DNA

**Collection:** During external ventricular drain (EVD) placement

**Sequencing:** Low-coverage whole genome sequencing (~0.5-1× coverage)

**Analysis Pipeline:**
1. Preprocessing: BWA-MEM alignment, quality control
2. CNV Detection: ichorCNA (tumor fraction estimation + copy number calling)
3. Subgroup Classification: CNV signature matching (WNT, SHH, Group 3, Group 4)
4. Risk Stratification: Schwalbe et al. 2017 criteria

**Reference Publications:**
- Liu et al. 2021, Cancer Cell (lcWGS for MB monitoring)
- Markowitz et al. 2025, NPJ Precision Oncology (Fragmentomics classification)
- Escudero et al. 2020, Nature Communications (CSF ctDNA subtyping)
- Schwalbe et al. 2017, Lancet Oncology (Risk stratification)

---

# LIMITATIONS AND DISCLAIMERS

1. **This is a research-use analysis** based on published methodologies. Results should be interpreted in conjunction with:
   - Histopathology
   - MRI imaging
   - Clinical presentation
   - Tissue-based molecular testing (methylation profiling preferred)

2. **Low tumor fraction** (<8%) may reduce classification accuracy

3. **CNV-based classification** does not detect:
   - Point mutations (e.g., TP53, CTNNB1)
   - Methylation patterns
   - Gene expression signatures

4. **Recommended confirmatory testing:**
   - Tissue-based DNA methylation profiling (gold standard)
   - Targeted sequencing for TP53, CTNNB1, and other driver mutations
   - Immunohistochemistry for subgroup markers

5. **Serial monitoring** with this assay may provide additional clinical value for:
   - Minimal residual disease detection
   - Early relapse prediction (3+ month lead time demonstrated)
   - Treatment response assessment

---

**Report Generated:** `r Sys.time()`  
**Pipeline Version:** MB-Classifier v1.0  
**Based on:** Liu 2021, Markowitz 2025, Escudero 2020

---

**For Clinical Questions Contact:**  
[Your Hospital Molecular Diagnostics Laboratory]
