---
title: "Medulloblastoma Classification Report"
output:
  pdf_document:
    toc: true
    toc_depth: 2
params:
  sample_id: ""
  classification_file: ""
  risk_file: ""
  ichorcna_dir: ""
  seg_file: ""
  fragmentomics_dir: ""
  fragmentomics_enabled: "false"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(data.table)
library(ggplot2)
library(knitr)
library(gridExtra)
```

# Sample Information

**Sample ID:** `r params$sample_id`  
**Analysis Date:** `r Sys.Date()`  
**Pipeline Version:** 2.0 with Fragmentomics Support

---

# Classification Results

```{r load_data}
# Load classification
classification <- fread(params$classification_file)

# Load risk
risk <- fread(params$risk_file)
```

## Molecular Subgroup

**Classification:** `r classification$classification[1]`  
**Confidence:** `r classification$confidence[1]`

### Rationale

```{r rationale}
rationale_text <- classification$rationale[1]
rationale_lines <- strsplit(rationale_text, "\\|")[[1]]
for (line in rationale_lines) {
  cat(paste0("- ", trimws(line), "\n"))
}
```

---

## Copy Number Alterations

```{r cnv_table}
cnv_results <- data.frame(
  Alteration = c("Monosomy 6", "Chromosome 9q loss", "MYC amplification", 
                 "Isochromosome 17q", "Chromosome 8 loss"),
  Status = c(
    ifelse(classification$monosomy_6[1], "DETECTED", "NOT DETECTED"),
    ifelse(classification$chr9q_loss[1], "DETECTED", "NOT DETECTED"),
    ifelse(classification$myc_amplification[1], "DETECTED", "NOT DETECTED"),
    ifelse(classification$isochromosome_17q[1], "DETECTED", "NOT DETECTED"),
    ifelse(classification$chromosome_8_loss[1], "DETECTED", "NOT DETECTED")
  ),
  Significance = c(
    "WNT signature (100%)",
    "SHH signature (32-100%)",
    "Group 3 signature (17%)",
    "Group 3/4 signature (42%/67%)",
    "Group 4 signature (20%)"
  )
)

kable(cnv_results, caption = "CNV Alterations Detected")
```

---

## Tumor Metrics

**Tumor Fraction:** `r sprintf("%.2f%%", classification$tumor_fraction[1] * 100)`  
**Ploidy:** `r sprintf("%.2f", classification$ploidy[1])`

Expected tumor fraction for CSF cfDNA: 8-72% (median 31%, Liu et al. 2021)

```{r tumor_fraction_plot}
tf <- classification$tumor_fraction[1]

ggplot(data.frame(x = c(0, 1)), aes(x)) +
  geom_rect(aes(xmin = 0.08, xmax = 0.72, ymin = 0, ymax = 1), 
            fill = "lightblue", alpha = 0.3) +
  geom_vline(xintercept = 0.31, linetype = "dashed", color = "blue", size = 1) +
  geom_vline(xintercept = tf, color = "red", size = 1.5) +
  annotate("text", x = 0.31, y = 0.9, label = "Median (31%)", angle = 90, vjust = -0.5) +
  annotate("text", x = tf, y = 0.9, label = sprintf("This sample (%.1f%%)", tf*100), 
           angle = 90, vjust = -0.5, color = "red") +
  labs(x = "Tumor Fraction", y = "", title = "Tumor Fraction Comparison") +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
```

---

# Fragmentomics Analysis

```{r fragmentomics_section, eval=(params$fragmentomics_enabled == "true")}
if (params$fragmentomics_enabled == "true") {
  
  cat("## Fragmentomics Results (Markowitz et al. 2025)\n\n")
  
  # Check if fragmentomics data available
  if ("fragmentomics_sl_ratio" %in% names(classification) && 
      !is.na(classification$fragmentomics_sl_ratio[1])) {
    
    sl_ratio <- classification$fragmentomics_sl_ratio[1]
    concordance <- classification$fragmentomics_concordance[1]
    
    cat(sprintf("**Short/Long Ratio:** %.3f\n\n", sl_ratio))
    cat(sprintf("**Literature Concordance:** %s\n\n", concordance))
    
    # Plot if available
    frag_plot <- file.path(params$fragmentomics_dir, 
                          paste0(params$sample_id, "_fragmentomics.png"))
    
    if (file.exists(frag_plot)) {
      cat("### Fragment Pattern Analysis\n\n")
      knitr::include_graphics(frag_plot)
    }
    
    cat("\n\n### Interpretation\n\n")
    
    if (concordance == "CONCORDANT") {
      cat("✓ Fragmentomics patterns are **CONCORDANT** with CNV-based classification.\n\n")
      cat("The S/L ratio and end motif patterns match published signatures for this MB subgroup.\n\n")
    } else if (concordance == "DISCORDANT") {
      cat("⚠ Fragmentomics patterns show **DISCORDANCE** with CNV classification.\n\n")
      cat("Further investigation may be warranted.\n\n")
    } else {
      cat("Fragmentomics analysis performed. Pattern analysis available above.\n\n")
    }
    
  } else {
    cat("Fragmentomics metrics not available in classification file.\n\n")
  }
  
} else {
  cat("## Fragmentomics Analysis\n\n")
  cat("Fragmentomics analysis was not performed for this sample.\n\n")
  cat("This is a CNV-only classification (Liu et al. 2021 approach).\n\n")
}
```

---

# ichorCNA Results

```{r ichorcna_plot}
# Include ichorCNA genome-wide plot
genomewide_plot <- file.path(params$ichorcna_dir, 
                             paste0(params$sample_id, "_genomeWide.pdf"))

if (file.exists(genomewide_plot)) {
  knitr::include_graphics(genomewide_plot)
} else {
  cat("ichorCNA plot not found.\n")
}
```

---

# Risk Stratification

```{r risk_info}
cat(sprintf("**Risk Group:** %s\n\n", risk$risk_group[1]))
cat(sprintf("**Metastasis:** %s\n\n", risk$metastasis[1]))
cat(sprintf("**Histology:** %s\n\n", risk$histology[1]))
cat(sprintf("**Risk Modifiers:** %s\n\n", risk$modifiers[1]))
```

## Important Notes

**This pipeline provides molecular subgroup classification only.**

Complete risk stratification requires integration of:

- Molecular subgroup (✓ provided by this pipeline)
- Tissue histology (large cell/anaplastic vs classic vs desmoplastic)
- MRI staging (M0 vs M+)
- Age at diagnosis
- Extent of resection
- Specific mutations (e.g., TP53 in SHH subgroup)

---

# Clinical Interpretation

```{r interpretation}
subgroup <- classification$classification[1]
confidence <- classification$confidence[1]

if (subgroup == "WNT") {
  cat("**WNT-activated medulloblastoma** has excellent prognosis with >95% survival.\n\n")
  cat("Treatment de-escalation may be considered to reduce long-term morbidity.\n\n")
  
} else if (subgroup == "SHH") {
  cat("**SHH-activated medulloblastoma** has variable prognosis depending on age and TP53 status.\n\n")
  cat("May be eligible for SMO inhibitor therapy (e.g., vismodegib).\n\n")
  
} else if (subgroup == "GROUP_3") {
  cat("**Group 3 medulloblastoma** often has poor prognosis, especially with MYC amplification.\n\n")
  cat("Aggressive treatment typically recommended.\n\n")
  
} else if (subgroup == "GROUP_4") {
  cat("**Group 4 medulloblastoma** has intermediate prognosis.\n\n")
  cat("Represents ~35% of all medulloblastomas.\n\n")
  
  if (params$fragmentomics_enabled == "true" && 
      "fragmentomics_concordance" %in% names(classification)) {
    if (classification$fragmentomics_concordance[1] == "CONCORDANT") {
      cat("Fragmentomics concordance provides additional validation of classification.\n\n")
    }
  }
}

cat("**Tumor fraction (", sprintf("%.1f%%", classification$tumor_fraction[1]*100), 
    ")** is within expected range for CSF cfDNA.\n\n", sep="")
```

---

# Methods

## CNV-based Classification

Classification follows Liu et al. 2021 methodology:

- Low-coverage whole genome sequencing (lcWGS)
- ichorCNA for CNV detection and tumor fraction estimation
- Hierarchical classification logic based on characteristic CNV signatures

CNV signatures from Escudero et al. 2020 (n=1,501 medulloblastomas).

```{r fragmentomics_methods, eval=(params$fragmentomics_enabled == "true")}
if (params$fragmentomics_enabled == "true") {
  cat("\n## Fragmentomics Analysis\n\n")
  cat("Fragment pattern analysis follows Markowitz et al. 2025 methodology:\n\n")
  cat("- Fragment length distribution (short: 90-150bp, long: 151-220bp)\n")
  cat("- Short-to-long (S/L) ratio calculation\n")
  cat("- 4bp fragment end motif analysis\n")
  cat("- Comparison to published MB subgroup-specific patterns\n\n")
  cat("Fragmentomics provides orthogonal validation of CNV-based classification.\n\n")
}
```

---

# References

1. **Liu et al. 2021** - Cancer Cell. "CSF liquid biopsy enables medulloblastoma classification"
2. **Escudero et al. 2020** - Acta Neuropathologica. "Molecular subgroup-specific CNV alterations"
3. **Schwalbe et al. 2017** - Lancet Oncology. "Clinical classification and outcome prediction"

```{r fragmentomics_ref, eval=(params$fragmentomics_enabled == "true")}
if (params$fragmentomics_enabled == "true") {
  cat("4. **Markowitz et al. 2025** - npj Precision Oncology. \"Genome-wide cfDNA fragmentation patterns\"\n")
}
```

---

**Report generated:** `r Sys.time()`
